Discrete Probability for Discrete Quantum Computing

### 2016-04-30 prob.tex
Add the cited document...

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware setting, it is more appropriate to
consider probability measures that map events to a finite set of
elements computable with a fixed set of resources. We will consider
two approaches: set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983} and interval-valued
probability measures~\cite{Dempster1967,Weichselberger2000,JamisonLodwick2004}.

@@

\begin{example}[Dempster-Shafer Theory of Evidence] We have five
  employees whose precise age is not known. All is given is a range for
  each employee:
\[\begin{array}{rcl}
\textrm{Age}(M_1) &=& [22,26] \\
\textrm{Age}(M_2) &=& [20,22] \\
\textrm{Age}(M_3) &=& [30,35] \\
\textrm{Age}(M_4) &=& [20,22] \\
\textrm{Age}(M_5) &=& [28,30] 
\end{array}\]
What is the probability that an employee's age is in the range
$[20,25]$? Looking at the data, it is \emph{possible} for $M_1$'s age
to be within the range, it is \emph{not possible} for $M_3$'s age to
be within the range, and it is \emph{certain} or \emph{necessary} that
$M_4$'s age is in the range. Clearly saying that an event is
\emph{necessary} is equivalent to saying that its complement is
\emph{not possible}. Aggregating the results for the five employees,
we can calculate that is necessary that 2 employees have their ages
within the range, and it is possible that 3 employees have their ages
within the range. We express this formally as saying that the
probability of the event is $[0.4,0.6]$. Now consider another event
asking whether the ages are some other disjoint range
$(25,40]$. Reasoning in a similar way we calculate that the
probability for this event is $[0.4,0.6]$. Now let's take the two events
together and ask about the possibility of the ages to be in the range
$[20,40]$. Clearly that probability must be $[1,1]$ as every
employee's age is in that range. The way to calculate this probability
from the individual probabilities of the events is as follows: 
\begin{itemize}
\item \amr{can you try to fix this... perhaps it is not consistent
    with the interval-valued formalization but it is really nice if we
    can make it work}
\end{itemize}
\qed\end{example}

>>>>

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware setting, it is more appropriate to
consider probability measures that map events to a finite set of
elements computable with a fixed set of resources. We will consider
two approaches: set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983} and interval-valued
probability
measures~\cite{Dempster1967,Zadeh1986,Weichselberger2000,JamisonLodwick2004}.

@@

\begin{example}[Dempster-Shafer Theory of Evidence] We have five
  employees whose precise age is not known. All is given is a range for
  each employee:
\[\begin{array}{rcl}
\textrm{Age}(M_1) &\in& [22,26] \\
\textrm{Age}(M_2) &\in& [20,22] \\
\textrm{Age}(M_3) &\in& [30,35] \\
\textrm{Age}(M_4) &\in& [20,22] \\
\textrm{Age}(M_5) &\in& [28,30] 
\end{array}\]
What is the probability that an employee's age is in the range
$[20,25]$? Looking at the data, it is \emph{possible} for $M_1$'s age
to be within the range, it is \emph{not possible} for $M_3$'s age to
be within the range, and it is \emph{certain} or \emph{necessary} that
$M_4$'s age is in the range. Clearly saying that an event is
\emph{necessary} is equivalent to saying that its complement is
\emph{not possible}. Aggregating the results for the five employees,
we can calculate that is necessary that 2 employees have their ages
within the range, and it is possible that 3 employees have their ages
within the range. We express this formally as saying that the
probability of the event is $[0.4,0.6]$. Now consider another event
asking whether the ages are some other disjoint range
$[26,40]$. Reasoning in a similar way we calculate that the
probability for this event is $[0.4,0.6]$. Now let's take the two events
together and ask about the possibility of the ages to be in the range
$[20,40]$. Clearly that probability must be $[1,1]$ as every
employee's age is in that range. The way to calculate this probability
from the individual probabilities of the events is as follows: 
\begin{itemize}
\item \amr{can you try to fix this... perhaps it is not consistent
    with the interval-valued formalization but it is really nice if we
    can make it work}
\end{itemize}
\qed\end{example}
